---
title: "Anatomical alteration after intensive PE with Ketamin/Midazolam agumantaion"
output: html_notebook
---

This notebook is extension to the jupiter notebook that extracts the results from TRACULA analysis

Load libraries 

```{r include=FALSE}
library("dplyr")
library("ggplot2")
library("lme4")
library("lmerTest")
library("emmeans")
library("tidyr")
library("rlist")
library("PerformanceAnalytics")
library("brms")
```

load WM

```{r}
setwd('/media/Data/Lab_Projects/KPE_PTSD_Project/other/Tracula')
WM      <-  read.table('KPE_tracula_Feb.csv', header = T, sep = ",")
```

Create database for analysis

```{r}
session <-  c(1, 2, 3)
tracts  <-  'rh.unc'
subs = c(1315, 1419)

WM <- subset(WM, (Tract %in% tracts) & (Ses %in% session))
WM <- subset(WM, !(Sub %in% subs))

# change data type for analysis
WM$med_cond  <- as.factor(WM$med_cond)
WM$Ses       <- as.factor(WM$Ses)

# add names for graphs
WM$DrugN[WM$med_cond==1] <- "Ketamine"
WM$DrugN[WM$med_cond==0] <- "Midazolam"

WM$Session[WM$Ses==1] <- "Day 1"
WM$Session[WM$Ses==2] <- "Day 7"
WM$Session[WM$Ses==3] <- "Followup"
```

Add a single motion parameter for WM analysis based on Yendiki et al., 2014

```{r}
m <- c(rep(0,2))
Q <- c(rep(0,4))
para <- c(rep(0,2))

m <- c(mean    (WM$AvgTranslation),    mean(WM$AvgRotation))
Q <- c(quantile(WM$AvgTranslation)[2], quantile(WM$AvgRotation)[2],
       quantile(WM$AvgTranslation)[4], quantile(WM$AvgRotation)[4])

para <- c(Q[3]-Q[1], Q[4]-Q[2])

for (i in 1:nrow(WM)){
  WM$TMI[i] <- sum((WM$AvgTranslation[i]-m[1])/para[1],(WM$AvgRotation[i]-m[2])/para[2])
}
  
```

Plots
```{r}
simple_model <- lmer(FA ~ gender + age + TMI + (1|Sub), data = WM)
WM$res <- resid(simple_model)

ggplot(WM, aes(x=Session, y=res, color=med_cond)) + geom_boxplot() + theme_minimal()  + theme(legend.position = "none") + facet_grid(~DrugN) + ylab("FA controlling for age, gender and motion")
```

Bayes model

```{r, include=FALSE}
bModel_p <- brm(FA ~ med_cond * Ses + gender + age + TMI + (1|Sub), data = WM, 
                family = student(), 
                prior = set_prior('normal(0, 1)'),
                iter = 2000, warmup = 1000, chains = 4, cores = 4, seed = 2019)
```

```{r}
print(summary(bModel_p, prob = .89) %>% print(digits = 4))
#print(coda::HPDinterval(as.mcmc(bModel_p, combine_chains = TRUE), prob = .89))
```

```{r}
plot(conditional_effects(bModel_p, prob = .89), ask = F)
```


```{r, include=FALSE}
Ket <- subset(WM, med_cond == 1)
Mid <- subset(WM, med_cond == 0)
KetModel_p <- brm(FA ~ Ses + gender + age + TMI + (1|Sub), data = Ket, family = student(), prior = set_prior('normal(0, 1)'),
              iter = 2000, warmup = 1000, chains = 4, cores = 4, seed = 2019)
MidModel_p <- brm(FA ~ Ses + gender + age + TMI + (1|Sub), data = Mid, family = student(), prior = set_prior('normal(0, 1)'),
              iter = 2000, warmup = 1000, chains = 4, cores = 4, seed = 2019)
```


```{r}
print(summary(KetModel_p, prob = .89) %>% print(digits = 4))
print(coda::HPDinterval(as.mcmc(KetModel_p, combine_chains = TRUE), prob = .89))

print(summary(MidModel_p, prob = .89) %>% print(digits = 4))
print(coda::HPDinterval(as.mcmc(MidModel_p, combine_chains = TRUE), prob = .89))

```



Function

```{r}
turn_to_wide <- function(db, columns, ses, db_name){
  sessions <- ses[ses!=1]
  return_col = c('Sub')
  wide <- db[columns]
  wide <- spread(wide, Ses, res)
  
  for(s in sessions){
    s_str <- toString(s)
    n = paste(db_name, s_str, '1', sep = '_')
    wide[n] <- wide[s_str] - wide['1']
    return_col <- list.append(return_col, n)
  }
  return (wide[return_col])
}
```

Create Delta's for correlation matrix

```{r}
columns <- c('Sub', 'Ses', 'res')

corr_db <- turn_to_wide(WM,   columns, session, 'rh_unc')
```

load functional files

```{r}
rest <- read.table('RS_conn_3Sessions.csv', header = T, sep = ",")
pcl <- read.table('pclitems_Feb.csv', header = T, sep = ",")
```

Create Deltas and select fields

```{r}
rest_columns <- c('amg_vmPFCant', 'hippAnt_vmPFCant')
end_col <- c('Sub')

sessions = c('2','3')
for (col in rest_columns){
  for (s in sessions){
    n <- paste(col, s, '1', sep = '_')
    n_s <- paste(col, s, sep = '')
    n_1 <- paste(col, '1', sep = '')
    rest[n] <- rest[n_s]-rest[n_1] 
    end_col <- list.append(end_col, n)
  }
}
rest <- rest[end_col]
```

Combine rest, task and anatomical data

```{r}
corr_db <- merge(merge(
              corr_db, 
              rest,   by = 'Sub'),
              pcl,   by = 'Sub')
```

Remove non numeric columns

```{r}
drop <- c('Sub', 'group', 'age')
'%ni%' <- Negate('%in%')
n <- names(corr_db)
n = n[n %ni% drop]

corr_db_clean <- corr_db[n]

sapply(corr_db_clean, class)
```

Correlation matrix
```{r}
part <- c('rh_unc_2_1', 'rh_unc_3_1', 'amg_vmPFCant_2_1', 'amg_vmPFCant_3_1', 'hippAnt_vmPFCant_2_1', 'hippAnt_vmPFCant_3_1', 'Re.experiencing_2', 'Re.experiencing_3')
unc <- corr_db_clean[part]

chart.Correlation(unc, histogram=TRUE, pch=19)
```

Bayes correlations

```{r}
bayes_corr <- function(x, y){
  newdata <- data_frame(x, y)
  bCorr <- brm(data = newdata, 
      family = student,
      mvbind(x, y) ~ 1,
      prior = c(prior(gamma(2, .1), class = nu),
                prior(normal(0, 1), class = Intercept),
                prior(normal(0, 1), class = sigma, resp = x),
                prior(normal(0, 1), class = sigma, resp = y),
                prior(lkj(1), class = rescor)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4, seed = 2015)
  return (bCorr)
}
```

Correlations between changes in re-experiencing symptoms and changes in RH UNC FA

```{r, include=FALSE}
reexp22 <- bayes_corr(corr_db_clean$rh_unc_2_1, corr_db_clean$Re.experiencing_2)
reexp23 <- bayes_corr(corr_db_clean$rh_unc_2_1, corr_db_clean$Re.experiencing_3)
reexp32 <- bayes_corr(corr_db_clean$rh_unc_3_1, corr_db_clean$Re.experiencing_2)
reexp33 <- bayes_corr(corr_db_clean$rh_unc_3_1, corr_db_clean$Re.experiencing_3)
```

```{r}
print("UNC post treatment - Reexp post treatment")
summary(reexp22, prob = .89)
print(coda::HPDinterval(as.mcmc(reexp22, combine_chains = TRUE), prob = .89))

print("UNC post treatment - Reexp follow up")
summary(reexp23, prob = .89)
print(coda::HPDinterval(as.mcmc(reexp23, combine_chains = TRUE), prob = .89))

print("UNC follow up - Reexp post treatment")
summary(reexp32, prob = .89)
print(coda::HPDinterval(as.mcmc(reexp32, combine_chains = TRUE), prob = .89))

print("UNC follow up - Reexp follow up")
summary(reexp33, prob = .89)
print(coda::HPDinterval(as.mcmc(reexp33, combine_chains = TRUE), prob = .89))

```

Correlations between changes in Amygdala vmPFC connectivity and changes in RH UNC FA

```{r, include=FALSE}
amy22 <- bayes_corr(corr_db_clean$rh_unc_2_1, corr_db_clean$amg_vmPFCant_2_1)
amy23 <- bayes_corr(corr_db_clean$rh_unc_2_1, corr_db_clean$amg_vmPFCant_3_1)
amy32 <- bayes_corr(corr_db_clean$rh_unc_3_1, corr_db_clean$amg_vmPFCant_2_1)
amy33 <- bayes_corr(corr_db_clean$rh_unc_3_1, corr_db_clean$amg_vmPFCant_3_1)
```

```{r}
print("UNC post treatment - Amygdala post treatment")
summary(amy22, prob = .89)
print(coda::HPDinterval(as.mcmc(amy22, combine_chains = TRUE), prob = .89))

print("UNC post treatment - Amygdala follow up")
summary(amy23, prob = .89)
print(coda::HPDinterval(as.mcmc(amy23, combine_chains = TRUE), prob = .89))

print("UNC follow up - Amygdala post treatment")
summary(amy32, prob = .89)
print(coda::HPDinterval(as.mcmc(amy32, combine_chains = TRUE), prob = .89))

print("UNC follow up - Amygdala follow up")
summary(amy33, prob = .89)
print(coda::HPDinterval(as.mcmc(amy33, combine_chains = TRUE), prob = .89))
```

Correlations between changes in Hiipocampus vmPFC connectivity and changes in RH UNC FA

```{r, include=FALSE}
hipp22 <- bayes_corr(corr_db_clean$rh_unc_2_1, corr_db_clean$hippAnt_vmPFCant_2_1)
hipp23 <- bayes_corr(corr_db_clean$rh_unc_2_1, corr_db_clean$hippAnt_vmPFCant_3_1)
hipp32 <- bayes_corr(corr_db_clean$rh_unc_3_1, corr_db_clean$hippAnt_vmPFCant_2_1)
hipp33 <- bayes_corr(corr_db_clean$rh_unc_3_1, corr_db_clean$hippAnt_vmPFCant_3_1)
```

```{r}
print("UNC post treatment - Hiippocampus post treatment")
summary(hipp22, prob = .89)
print(coda::HPDinterval(as.mcmc(hipp22, combine_chains = TRUE), prob = .89))

print("UNC post treatment - Hiippocampus follow up")
summary(hipp23, prob = .89)
print(coda::HPDinterval(as.mcmc(hipp23, combine_chains = TRUE), prob = .89))

print("UNC follow up - Hiippocampus post treatment")
summary(hipp32, prob = .89)
print(coda::HPDinterval(as.mcmc(hipp32, combine_chains = TRUE), prob = .89))

print("UNC follow up - Amygdala follow up")
summary(hipp33, prob = .89)
print(coda::HPDinterval(as.mcmc(hipp33, combine_chains = TRUE), prob = .89))

```

# plotting the correlations

```{r}
library(reshape2)
colC <- c('amg_vmPFCant_2_1', 'amg_vmPFCant_3_1', 'hippAnt_vmPFCant_2_1', 'hippAnt_vmPFCant_3_1') 

long <- melt(corr_db_clean, id.vars=c("rh_unc_2_1", "rh_unc_3_1"))

connectivety <- subset(long, variable %in% colC)

connect <- separate(data = connectivety, col = variable, into = c("ROI", "frontal", 'session', 'clear'), sep = "_")
```

```{r}
ROI.labs <- c('Amygdala', 'Hippocampus')
names(ROI.labs) <- c('amg', 'hippAnt')

session.labs <- c('Post treatment', 'Follow up')
names(session.labs) <- c('2', '3')

rho_text <- tibble('label' = c('ρ=0.19', 'ρ=0.23', 'ρ=0.40*', 'ρ=0.25'), 
                   ROI = c('amg','amg','hippAnt','hippAnt'),
                   session = c('2','3','2','3'))

ggplot(connect, aes(rh_unc_2_1, value)) + geom_point() + 
  geom_smooth(level = 0.89, method = "lm") +# theme_minimal() +
  xlab('WMI change Post treatment compared to baseline') + ylab('Prefortnal connectivety changes compared to baseline') +
  facet_grid(ROI ~ session, labeller = labeller(session = session.labs, ROI = ROI.labs)) + 
  geom_text(size = 5, data = rho_text, 
            mapping = aes(x = -0.02, y = Inf, label = label),
            hjust   = 1.05,
            vjust   = 1.5)
```

```{r}
rho_text <- tibble('label' = c('ρ=0.15', 'ρ=0.32#', 'ρ=0.54*', 'ρ=0.16'), 
                   ROI = c('amg','amg','hippAnt','hippAnt'),
                   session = c('2','3','2','3'))

ggplot(connect, aes(rh_unc_3_1, value)) + geom_point() + 
  geom_smooth(level = 0.89, method = "lm") + # theme_minimal() +
  xlab('WMI change follow up compared to baseline') + ylab('Prefortnal connectivety changes compared to baseline') +
  facet_grid(ROI ~ session, labeller = labeller(session = session.labs, ROI = ROI.labs)) +
  geom_text(size = 5, data = rho_text, 
            mapping = aes(x = -0.02, y = Inf, label = label),
            hjust   = 1.05,
            vjust   = 1.5)
```
```{r}
colR <- c('Re.experiencing_2', 'Re.experiencing_3')

reexp <- subset(long, variable %in% colR)
reexp <- separate(data = reexp,  col = variable, into = c('symptom', 'treatment_session'), sep = "_")
names(reexp)[names(reexp) == "value"] <- "caps"

ree <- melt(reexp, id.vars=c('symptom', 'treatment_session','caps'))
ree <- separate(data = ree,  col = variable, into = c('hem','tract', 'tract_session','clear'), sep = "_")
```

```{r}

tract_session.labs <- c('Post treatment', 'Follow up')
treatment_session.labs <- c('Post treatment', 'Follow up')
names(treatment_session.labs) <- c('2', '3')
names(tract_session.labs) <- c('2', '3')

rho_text <- tibble('label' = c('ρ=0.45*', 'ρ=0.41*', 'ρ=0.29', 'ρ=0.3#'), 
                   treatment_session = c('2','2','3','3'),
                   tract_session = c('2','3','2','3'))

ggplot(ree, aes(value, caps)) + geom_point() + 
  geom_smooth(level = 0.89, method = "lm") + #theme_minimal() +
  xlab('WMI changes comapred to base line') + ylab('Reexperiencing sypotom change compared to baseline') +
  facet_grid(tract_session ~ treatment_session, 
             labeller = labeller(tract_session = tract_session.labs, treatment_session = treatment_session.labs)) + 
  geom_text(size = 5, data = rho_text, 
            mapping = aes(x = -0.02, y = Inf, label = label),
            hjust   = 1.05,
            vjust   = 1.5)
```


```{r}
hipp_unc <- corr_db_clean[c("rh_unc_2_1","rh_unc_3_1", "hippAnt_vmPFCant_2_1","hippAnt_vmPFCant_3_1")]
hipp_unc['Sub'] <- 1:24
hipp_unc1 <- melt(hipp_unc, id.vars = 'Sub', measure.vars = c("rh_unc_2_1","rh_unc_3_1"), variable.name = "wm_ses", value.name = 'unc')
hipp_unc1 <- rbind(hipp_unc1, hipp_unc1)
hipp_unc2 <- melt(hipp_unc, id.vars = 'Sub', measure.vars = c("hippAnt_vmPFCant_2_1","hippAnt_vmPFCant_3_1"), variable.name = "fc_ses", value.name = 'fc')
hipp_unc2 <- rbind(subset(hipp_unc2, fc_ses == 'hippAnt_vmPFCant_2_1'),
                   hipp_unc2, 
                   subset(hipp_unc2, fc_ses == 'hippAnt_vmPFCant_3_1')) 
hipp_unc <- cbind(hipp_unc1,hipp_unc2)
hipp_unc <- hipp_unc[-4]
```

```{r}
ses.labs <- c('rsFC Post treatment', 'rsFC Follow up','WMI Post treatment', 'WMI Follow up')
names(ses.labs) <- c('hippAnt_vmPFCant_2_1', 'hippAnt_vmPFCant_3_1','rh_unc_2_1', 'rh_unc_3_1')

rho_text <- tibble('label' = c('ρ=0.41*', 'ρ=0.54*', 'ρ=0.25', 'ρ=0.16'),
                   fc_ses = c('hippAnt_vmPFCant_2_1','hippAnt_vmPFCant_2_1','hippAnt_vmPFCant_3_1','hippAnt_vmPFCant_3_1'),
                   wm_ses = c('rh_unc_2_1','rh_unc_3_1','rh_unc_2_1','rh_unc_3_1'))

ggplot(hipp_unc, aes(unc, fc)) + geom_point() + 
  geom_smooth(level = 0.89, method = "lm") + #theme_minimal() +
  xlab('WMI changes comapred to base line') + ylab('Functional connectivety changes compared to baseline') +
  facet_grid(fc_ses ~ wm_ses,
            labeller = as_labeller(ses.labs)) + 
  geom_text(size = 5, data = rho_text,
            mapping = aes(x = -0.02, y = Inf, label = label),
            hjust   = 1.05,
            vjust   = 1.5)
```

